---
title: "README_Zack"
author: "Zack Grzywacz"
date: "October 23, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
loblolly <- Loblolly
```

1. The dataset contains 84 observations of 3 variables.


```{r}
str(loblolly$height)
typeof(loblolly$height)
```

```{r}
str(loblolly$age)
typeof(loblolly$age)
```

```{r}
str(loblolly$Seed)
typeof(loblolly$Seed)
```

2. Height and age are the "double" data type, while Seed is the type "integer". However, while height and age are numeric, Seed is an ordered factor. 

3. To convert Seed to an integer, we could use the following command:

```{r}
loblolly$Seed <- as.integer (loblolly$Seed)
```

```{r}
str(loblolly$Seed)
typeof(loblolly$Seed)
```

4. To record the date in the computer as a column:

```{r}
loblolly$date <- Sys.Date()
head(loblolly)
```

5. To add the 'mature' vector, we will use a for loop to create a column of 1s and 0s. Then, we will coerce it to logical:

```{r}
for (yr in 1:84)
{
  if (loblolly$age[yr] > 9){
    loblolly$mature[yr] <- 1
    }
  else{
    loblolly$mature[yr] <- 0
  }
}
loblolly$mature <- as.logical(loblolly$mature)
```

6. To create the list:

```{r}
list_example <- list(c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"), month.abb, 1:31)
list_example
```

7. We will create the matrix. Then, we will add row names and column names:

```{r}
my_matrix <- matrix(0, ncol=5, nrow=4, data=seq(1:5))
colnames(my_matrix) <- c("Dory","Edna","Eva","Boo","Violet")
rownames(my_matrix) <- c("FindingNemo","Incredibles","Wall-E","MonstersInc")
my_matrix
```

8. 

First we must remove Violet from the matrix:

```{r}
my_matrix <- my_matrix[,-5]
my_matrix
```

```{r}
for (data in my_matrix){
if(my_matrix[data] > 1){
  my_matrix[data] = 0
}
}
```


```{r}
class(my_matrix) <- "logical"
my_matrix
```